---
title: "Lab11"
author: "Evelyn Zhang"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exploratory Data Analysis
```{r}
data <- read.csv("./class11_LAB_dataFrame_20190930T2158.csv")
data
plot(data$x,data$y)
plot(data$a,data$y)
plot(data$b,data$y)
plot(data$c,data$y)
# As x increases, y increases, but the relationship is nonlinear. 
# There is no obvious relationship between y and a
# y vs b resembles a quadratic relationship
# As c values increase, y values become more spread out and as c values increase more, y values become closer together again
```

#### Regression
```{r}
model1 <- lm(y~a,data=data)
plot(data$a,data$y)
lines(data$a,predict.lm(model1,data.frame(a=data$a)),col='red')
# y values are concentrated around low a values but there is no obvious relationship shown in the SLR model. 
```

#### Transforms
```{r}
options(repr.plot.width=7*1, repr.plot.height=7)
par(mfrow=c(2,2))
transformedX = log(data$a)
plot(transformedX,data$y,xlab='log(a)',ylab='Y',tck = 0.02)
model <- lm(y~log(a),data=data)

predictions <- predict.lm(model,data.frame(a=data$a))
lines(transformedX
      ,predictions
      ,col='red')

transformedX = (data$a)^0.5
plot(data$a^0.5,data$y,xlab='a^0.5',ylab='Y',tck = 0.02)
model <- lm(y~I(a^0.5),data=data)

predictions <- predict.lm(model,data.frame(a=data$a))
lines(transformedX
      ,predictions
      ,col='red')

transformedX = (data$a)^0.333
plot(data$a^0.333,data$y,xlab='a^1/3',ylab='Y',tck = 0.02)
model <- lm(y~I(a^0.333),data=data)

predictions <- predict.lm(model,data.frame(a=data$a))
lines(transformedX
      ,predictions
      ,col='red')

transformedX = (data$a)^0.25
plot(data$a^0.25,data$y,xlab='a^0.25',ylab='Y',tck = 0.02)
model <- lm(y~I(a^0.25),data=data)

predictions <- predict.lm(model,data.frame(a=data$a))
lines(transformedX
      ,predictions
      ,col='red')

#The relasionship between y and a appears more linear. 
```


#### Residuals
```{r}
options(repr.plot.width=8, repr.plot.height=8)
par(mfrow=c(2,2))

modelLog <- lm(y~log(a),data=data)
epsilonsLog <- residuals(modelLog)
hist(epsilonsLog,breaks=seq(-20,30,2),col='green',,main="log(a)")

modelHalf <- lm(y~I(a^0.5),data=data)
epsilonsHalf <- residuals(modelHalf)
hist(epsilonsHalf,breaks=seq(-20,30,2),col='green',main="a^0.50")


modelThird <- lm(y~I(a^0.33),data=data)
epsilonsThird <- residuals(modelThird)
hist(epsilonsThird,breaks=seq(-20,30,2),col='green',main="a^0.33")


modelQuarter <- lm(y~I(a^0.25),data=data)
epsilonsQuarter <- residuals(modelQuarter)
hist(epsilonsQuarter,breaks=seq(-20,30,2),col='green',main="a^0.25")

```

#### QQ
```{r}
par(mfrow=c(2,2))

# plot 1
epsilonsLog = sort(epsilonsLog)
N = length(epsilonsLog)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsLog
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsLog~theoreticalNormal))
title('log(a)')


# plot 2
epsilonsHalf = sort(epsilonsHalf)
N = length(epsilonsHalf)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsHalf
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsHalf~theoreticalNormal))
title('a^0.5')


# plot 3
epsilonsThird = sort(epsilonsThird)
N = length(epsilonsThird)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsThird
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsThird~theoreticalNormal))
title('a^0.33')

# plot 4
epsilonsQuarter = sort(epsilonsQuarter)
N = length(epsilonsQuarter)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsQuarter
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsQuarter~theoreticalNormal))
title('a^0.25')

#The residuals and qq plot are very similar between the four models. 
```

#### MLR

```{r}
modelMLR <- lm(data$y ~ data$x + data$a + data$b + data$c, data=data)

options(repr.plot.width=8, repr.plot.height=8)
par(mfrow=c(2,2))

epsilons <- residuals(modelMLR)
hist(epsilons,breaks=seq(-20,30,2),col='green',,main="MLR")

epsilons = sort(epsilons)
N = length(epsilons)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilons
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilons~theoreticalNormal))
title('MLR')
```

#1. Does your MLR satisfy the LINE assumptions?
# No

#2. If not, what transforms do you recommend be applied to each explanatory variable and why?
# Transform x to e^x and b to b^2 because the plots look so.

#3. Transform the variable you feel most violates the LINE assumptions.

```{r}
modelMLR <- lm(data$y ~ exp(data$x) + data$a + data$b + data$c, data=data)

options(repr.plot.width=8, repr.plot.height=8)
par(mfrow=c(2,2))

epsilons <- residuals(modelMLR)
hist(epsilons,breaks=seq(-20,30,2),col='green',,main="MLR")

epsilons = sort(epsilons)
N = length(epsilons)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilons
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilons~theoreticalNormal))
title('MLR')
```

#4. Does transforming this variable help? why or why not?
# No. Other variables may need to be transformed as well. 


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
