---
title: "Lab11_Sec02"
author: "Susan Khan"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LAB11

## Data Transforms

1. Read the Dataset "class11\_LAB\_dataFrame\_20190930T2158.csv" into a dataframe.
2. The target variable is **y**

```{r}
library(readr)
DataLab11 <- read_csv("class11_LAB_dataFrame_20190930T2158.csv")


```
#### Exploratory Data Analysis
1. Plot y vs.
   1. x
   2. a
   3. b
   4. c
   
```{r}

plot(DataLab11$x,DataLab11$y)
plot(DataLab11$a, DataLab11$y)
plot(DataLab11$b, DataLab11$y)
plot(DataLab11$c, DataLab11$y)



```

Summarize the relationships between $y$ and the set of explanatory variables (x,a,b,c).

```{r}
# For relationship between x and y appears non linear and thus breaks the Linearity assumption.

```

#### Regression
1. Please fit a simple linear regression model between **y** and **a**
2. Make a scatter plot of **y** and **a**
3. Overlay the regression line
4. Comment on the predicted model versus ground truth

```{r}
model1= lm(y~a, data = DataLab11)
plot(DataLab11$a, DataLab11$y)
abline (model1)

#There appears to be a very slight positive relationship between a and y. However, in reality all the data points are mostly clustured between 1.5 and 1.6. This breaks our equal variance assumption.

```

#### Transforms
1. Transform **a** 4 different ways
2. Fit a linear regression
3. Overlay the regression line
4. Comment on the predicted model versus ground truth

```{r}
options(repr.plot.width=7*1, repr.plot.height=7)
par(mfrow=c(2,2))
transformedA = log(DataLab11$a)
plot(transformedA,DataLab11$y,xlab='log(a)',ylab='Y',tck = 0.02)
model <- lm(y~log(a),data=DataLab11)

predictions <- predict.lm(model,data.frame(a=DataLab11$a))
lines(transformedA
      ,predictions
      ,col='red')

transformedA = (DataLab11$a)^0.5
plot(DataLab11$a^0.5,DataLab11$y,xlab='a^0.5',ylab='Y',tck = 0.02)
model <- lm(y~I(a^0.5),data=DataLab11)

predictions <- predict.lm(model,data.frame(a=DataLab11$a))
lines(transformedA
      ,predictions
      ,col='red')

transformedA = (DataLab11$a)^0.333
plot(DataLab11$a^0.333,DataLab11$y,xlab='a^1/3',ylab='Y',tck = 0.02)
model <- lm(y~I(a^0.333),data=DataLab11)

predictions <- predict.lm(model,data.frame(a=DataLab11$a))
lines(transformedA
      ,predictions
      ,col='red')

transformedA = (DataLab11$a)^0.25
plot(DataLab11$a^0.25,DataLab11$y,xlab='a^0.25',ylab='Y',tck = 0.02)
model <- lm(y~I(a^0.25),data=DataLab11)

predictions <- predict.lm(model,data.frame(a=DataLab11$a))
lines(transformedA
      ,predictions
      ,col='red')


#there appears to be a slightly positive relationship between a and y even after transforming the variable a in 4 different ways.



```

#### Residuals
1. Plot a histogram of the above 4 model residuals

```{r}
options(repr.plot.width=8, repr.plot.height=8)
par(mfrow=c(2,2))

modelLog <- lm(y~log(a),data=DataLab11)
epsilonsLog <- residuals(modelLog)
hist(epsilonsLog,col='green', main="log(a)")

modelHalf <- lm(y~I(a^0.5),data=DataLab11)
epsilonsHalf <- residuals(modelHalf)
hist(epsilonsHalf, col='green',main="a^0.50")


modelThird <- lm(y~I(a^0.33),data=DataLab11)
epsilonsThird <- residuals(modelThird)
hist(epsilonsThird,col='green',main="a^0.33")


modelQuarter <- lm(y~I(a^0.25),data=DataLab11)
epsilonsQuarter <- residuals(modelQuarter)
hist(epsilonsQuarter, col='green',main="a^0.25")
```

#### QQ
1. Plot a qqplot of the above 4 model residuals

```{r}
par(mfrow=c(2,2))

# plot 1
epsilonsLog = sort(epsilonsLog)
N = length(epsilonsLog)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsLog
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsLog~theoreticalNormal))
title('log')

# plot 2
epsilonsHalf = sort(epsilonsHalf)
N = length(epsilonsHalf)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsHalf
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsHalf~theoreticalNormal))
title('x^1/2')


# plot 3
epsilonsThird = sort(epsilonsThird)
N = length(epsilonsThird)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsThird
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsThird~theoreticalNormal))


# plot 4
epsilonsQuarter = sort(epsilonsQuarter)
N = length(epsilonsQuarter)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsQuarter
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Normal')
abline(lm(epsilonsQuarter~theoreticalNormal))

```

How do the residuals and qq plot compare between your 4 different transforms?

```{r}
# write a comment

#After computing the residual and qq plot of the 4 models, it appears that none of the data transformation satisfies the normality assumption.

```

#### MLR
1. Please fit a multiple linear regression model between **y** and (**x**, **a**, **b**, **c**)
2. Plot a histogram of the above 4 model residuals
3. Plot a qqplot of the above 4 model residuals

```{r}
par(mfrow=c(2,2))

MLRmodellog= lm(y~log(a)+x+b+c, data = DataLab11)
epsilonsLog1 <- residuals(MLRmodellog)
hist(epsilonsLog1, col='yellow',main="log(a)")


MLRmodelhalf= lm(y~I(a^0.5)+x+b+c, data = DataLab11)
epsilonshalf1 <- residuals(MLRmodelhalf)
hist(epsilonshalf1, col='yellow',main="(a^0.5)")


MLRmodelthird= lm(y~I(a^0.33)+x+b+c, data = DataLab11)
epsilonsthird1 <- residuals(MLRmodelthird)
hist(epsilonsthird1, col='yellow',main="(a^0.33)")

MLRmodelquarter= lm(y~I(a^0.25)+x+b+c, data = DataLab11)
epsilonsquarter1 <- residuals(MLRmodelquarter)
hist(epsilonsquarter1, col='yellow',main="(a^0.25)")




```

```{r}

par(mfrow=c(2,2))

# plot 1
epsilonsLog1 = sort(epsilonsLog1)
N = length(epsilonsLog1)
probs = seq(1,N)/(N+1)
theoreticalNormal1 = qnorm(probs)

plot(theoreticalNormal1,epsilonsLog1
     ,xlab='Theoretical Normal1'
     ,ylab='Empirical Normal1')
abline(lm(epsilonsLog1~theoreticalNormal1))
title('log')

# plot 2
epsilonshalf1 = sort(epsilonshalf1)
N = length(epsilonshalf1)
probs = seq(1,N)/(N+1)
theoreticalNormal1 = qnorm(probs)

plot(theoreticalNormal1,epsilonshalf1
     ,xlab='Theoretical Normal1'
     ,ylab='Empirical Normal1')
abline(lm(epsilonshalf1~theoreticalNormal1))
title('a^1/2')


# plot 3
epsilonsthird1 = sort(epsilonsthird1)
N = length(epsilonsthird1)
probs = seq(1,N)/(N+1)
theoreticalNormal1 = qnorm(probs)

plot(theoreticalNormal1,epsilonsthird1
     ,xlab='Theoretical Normal1'
     ,ylab='Empirical Normal1')
abline(lm(epsilonsthird1~theoreticalNormal1))
title('a^1/3')

# plot 4
epsilonsquarter1 = sort(epsilonsquarter1)
N = length(epsilonsquarter1)
probs = seq(1,N)/(N+1)
theoreticalNormal1 = qnorm(probs)

plot(theoreticalNormal1,epsilonsquarter1
     ,xlab='Theoretical Normal1'
     ,ylab='Empirical Normal1')
abline(lm(epsilonsquarter1~theoreticalNormal1))
title('a^1/4')

```


1. Does your MLR satisfy the LINE assumptions?
2. If not, what transforms do you recommend be applied to each explanatory variable and why?
3. Transform the variable you feel most violates the LINE assumptions.
4. Does transforming this variable help? why or why not?

```{r}

MLRmodel= lm(y~a+x+b+c, data = DataLab11)
plot(MLRmodel)
#After computing the multiple linear regression including variables b, c and d, our model does not change at all. It does not meet the equal variance assumption because we see from the residual vs fitted plot that there is a clear pattern of the residuals. Additionally, from the normal q-q plot, the MLR model fails the normality assumption.

MLRmodelTransform2= lm(y~I(x^0.5)+ log(a)+b+c, data = DataLab11)
plot(MLRmodellog2)

#After looking at the scatterplots above, I chose to transform x to x^0.5 and executed a log transformation on a. The results show that the residuals are spread across more evenly and the q-q plot shows that this transformed model fits the normality assumption much better.

```

