## Data Transforms
Name: Kaiwen Lu

1. Read the Dataset "class11\_LAB\_dataFrame\_20190930T2158.csv" into a dataframe.
2. The target variable is **y**
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(GGally) 
library(gridExtra)
data<-read.csv('class11_LAB_dataFrame_20190930T2158.csv')
data<-data%>%select(x,a,b,c,y)
```
#### Exploratory Data Analysis
1. Plot y vs.
   1. x
   2. a
   3. b
   4. c
```{r}
ggplot(data=data,mapping=aes(x=x,y=y) )+geom_point()
ggplot(data=data,mapping=aes(x=a,y=y) )+geom_point()
ggplot(data=data,mapping=aes(x=b,y=y) )+geom_point()
ggplot(data=data,mapping=aes(x=c,y=y) )+geom_point()

```

#### Regression
1. Please fit a simple linear regression model between **y** and **a**
2. Make a scatter plot of **y** and **a**
3. Overlay the regression line
4. Comment on the predicted model versus ground truth
The linear regression fits well for x values greater than 1.65 but under estimate y values for x values less than 1/2. The relationship between x and y appears non-linear. It breaks the L assumption. 
```{r}
model_a<-lm(y~a, data=data)
summary(model_a)
ggplot(data=data,mapping=aes(x=a,y=y) )+geom_point()+geom_smooth()


```

#### Transforms
1. Transform **a** 4 different ways
2. Fit a linear regression
3. Overlay the regression line
4. Comment on the predicted model versus ground truth
There is some curveness in the fitted line and the transformation doesn't help much. 

```{r}
#log a
logdata<-data %>% mutate(
  log_a=log(a)
  )
log<-lm(y~log_a, data=logdata)
ggplot(data=logdata,mapping=aes(x=log_a,y=y) )+geom_point()+geom_smooth()

#sqrt a
sqrtdata<-data%>% mutate(
  sqrt_a=sqrt(a)
)
sqrt<-lm(y~sqrt_a,data=sqrtdata)
ggplot(data=sqrtdata,mapping=aes(x=sqrt_a,y=y) )+geom_point()+geom_smooth()
# 
# #square a
# sqdata<-data%>% mutate(
#   sq_a=a^2
# )
# sq<-lm(y~sq_a,data=sqdata)
# ggplot(data=sqdata,mapping=aes(x=sq_a,y=y) )+geom_point()+geom_smooth()

#sqrt log a
sqrtlogdata<-data %>% mutate(
  sqrtlog_a=sqrt(log(a))
  )
sqrtlog<-lm(y~sqrtlog_a, data=sqrtlogdata)
ggplot(data=sqrtlogdata,mapping=aes(x=sqrtlog_a,y=y) )+geom_point()+geom_smooth()

#log log a
loglogdata<-data %>% mutate(
  loglog_a=log(log(a))
  )
loglog<-lm(y~loglog_a, data=loglogdata)
ggplot(data=loglogdata,mapping=aes(x=loglog_a,y=y) )+geom_point()+geom_smooth()




```

#### Residuals
1. Plot a histogram of the above 4 model residuals

```{r}

logdata <- logdata %>%
mutate(
residuals1 = residuals(log),
fitted = predict(log)
)
ggplot(data = logdata, mapping = aes(x = residuals1)) +
geom_histogram() +
ggtitle("Residuals")

sqrtdata <- sqrtdata %>%
mutate(
residuals2 = residuals(sqrt),
fitted = predict(sqrt)
)
ggplot(data = sqrtdata , mapping = aes(x = residuals2)) +
geom_histogram() +
ggtitle("Residuals")

sqrtlogdata <- sqrtlogdata %>%
mutate(
residuals3 = residuals(sqrtlog),
fitted = predict(sqrtlog)
)
ggplot(data = sqrtlogdata, mapping = aes(x = residuals3)) +
geom_histogram() +
ggtitle("Residuals")

loglogdata <- loglogdata %>%
mutate(
residuals4 = residuals(loglog),
fitted = predict(loglog)
)
ggplot(data = loglogdata, mapping = aes(x = residuals4)) +
geom_histogram() +
ggtitle("Residuals")


```


#### QQ
How do the residuals and qq plot compare between your 4 different transforms?
The residuals and qq plots between my 4 different transformations are very similar. 
```{r}
ggplot(data = logdata, mapping = aes(sample = residuals1)) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

ggplot(data = sqrtdata, mapping = aes(sample = residuals2)) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

ggplot(data = sqrtlogdata, mapping = aes(sample = residuals3)) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

ggplot(data = loglogdata, mapping = aes(sample = residuals4)) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")
```



#### MLR
1. Does your MLR satisfy the LINE assumptions?
(1) Linearity: The scatter plots between y and x,y,c all seem to violate the linearity. Transformation is needed.
(2)Independence: We assume the observations are independent. 
(3)Normal distributions of residuals: From the residuals plots and QQ plots, the models are approximately normally distributed.
(4)Equal standard deviation: The standard deviation of the explanatory variables don't look equal from the plots. 

2. If not, what transforms do you recommend be applied to each explanatory variable and why?
For variable x, try exponentiate the x; For a and d, they look approximately linear so transformation does not help much. For c, try replacing x^2 with z since it looks like a parabola model. So that y=x^2=z. 

3. Transform the variable you feel most violates the LINE assumptions.
Transform x. 

4. Does transforming this variable help? why or why not?
Transforming this variable helps.The relationship between the transformed $x$ variable and $y$ appears linear and we would restore our L assumption.

```{r}
#Multiple linear regression model 
MLR<-lm(y~x+a+b+c,data=data)
summary(MLR)

#Residuals 
#Histogram of the above 4 model residuals 
model_x<-lm(y~x,data=data)
ggplot(data =data, mapping = aes(x =residuals(model_x))) +
geom_histogram() +
ggtitle("Residuals")

ggplot(data = data, mapping = aes(x = residuals(model_a))) +
geom_histogram() +
ggtitle("Residuals")

model_b<-lm(y~b,data=data)
ggplot(data = data, mapping = aes(x = residuals(model_b))) +
geom_histogram() +
ggtitle("Residuals")

model_c<-lm(y~c,data=data)
ggplot(data = data, mapping = aes(x = residuals(model_c))) +
geom_histogram() +
ggtitle("Residuals")

#A qqplot of the above 4 model residuals
ggplot(data = data, mapping = aes(sample = residuals(model_x))) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

ggplot(data = data, mapping = aes(sample = residuals(model_a))) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

ggplot(data = data, mapping = aes(sample = residuals(model_b))) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

ggplot(data = data, mapping = aes(sample = residuals(model_c))) +
stat_qq() +
stat_qq_line() +
ggtitle("Residuals Q-Q")

#Equal variane 
plot(data$x,residuals(model_x), tck=0.01)
title('x')

plot(data$a, residuals(model_a),tck=0.01)
title('a')

plot(data$b, residuals(model_b),tck=0.01)
title('b')

plot(data$c, residuals(model_c),tck=0.01)
title('c')



#Plot of MLR
ggpairs(data)

#Transform the variable x
ggplot(aes(x=exp(x),y=y),data=data)+geom_point()




```

