# LAB11
title: "Lab11"
author: "Yue Qiu"
date: "10/11/2019"


## Data Transforms

1. Read the Dataset "class11\_LAB\_dataFrame\_20190930T2158.csv" into a dataframe.
2. The target variable is **y**

```{r}
data <- read.csv("class11_LAB_dataFrame_20190930T2158.csv")
```

#### Exploratory Data Analysis
1. Plot y vs.
   1. x
   2. a
   3. b
   4. c
*Summarize the relationships between $y$ and the set of explanatory variables (x,a,b,c).

```{r}
plot(data$x,data$y)
model1 <- lm(y~x,data=data)
lines(data$x,predict.lm(model1,data.frame(x=data$x)),col='red')
summary(model1)
```
The relationship between y and x is curvy and points are going upward.

```{r}
plot(data$a,data$y)
model2 <- lm(y~a,data=data)
lines(data$a,predict.lm(model2,data.frame(a=data$a)),col='red')
summary(model2)
```
Most points are clustered on the left side, but overall the trend is going to the right side. Their relationship is linear but flat


```{r}
plot(data$b,data$y)
model3 <- lm(y~b,data=data)
lines(data$c,predict.lm(model3,data.frame(b=data$b)),col='red')
summary(model3)
```
The relationship is slightly curved down. Points are spread out and it seems to have many outliers. Weaker sign of linear relationship.

```{r}
plot(data$c,data$y)
model4 <- lm(y~c,data=data)
lines(data$c,predict.lm(model4,data.frame(c=data$c)),col='red')
summary(model4)
```
The relationship between y and c is lienar but flat. Points are spread out.


#### Regression 
1. Please fit a simple linear regression model between **y** and **a**
2. Make a scatter plot of **y** and **a**
3. Overlay the regression line
4. Comment on the predicted model versus ground truth
```{r}
plot(data$a,data$y)
model2 <- lm(y~a,data=data)
lines(data$a,predict.lm(model2,data.frame(a=data$a)),col='red')
summary(model2)
```
The predicted model is y=6.195a-4.641
The distribution of this plot is not healthy because points are clustered on the left and may have outliers. So the model does not fit with the data of a.


#### Transforms
1. Transform **a** 4 different ways
2. Fit a linear regression
3. Overlay the regression line
4. Comment on the predicted model versus ground truth

```{r}
transformedA = log(data$a)
plot(transformedA,data$y,xlab='log(a)',ylab='Y',tck = 0.02)
model1 <- lm(y~log(a),data=data)
predictions <- predict.lm(model1,data.frame(a=data$a))
lines(transformedA
      ,predictions
      ,col='red')
summary(model1)
```

```{r}
transformedA = (data$x)^0.5
plot(data$a^0.5,data$y,xlab='a^0.5',ylab='Y',tck = 0.02)
model2 <- lm(y~I(a^0.5),data=data)
predictions <- predict.lm(model2,data.frame(a=data$a))
lines(transformedA
      ,predictions
      ,col='red')
summary(model2)
```

```{r}
transformedA = (data$a)^0.333
plot(data$a^0.333,data$y,xlab='a^1/3',ylab='Y',tck = 0.02)
model3 <- lm(y~I(a^0.333),data=data)
predictions <- predict.lm(model3,data.frame(a=data$a))
lines(transformedA
      ,predictions
      ,col='red')
summary(model3)
```

```{r}
transformedA = (data$a)^0.25
plot(data$a^0.25,data$y,xlab='a^0.25',ylab='Y',tck = 0.02)
model4 <- lm(y~I(a^0.25),data=data)
predictions <- predict.lm(model4,data.frame(a=data$a))
lines(transformedA
      ,predictions
      ,col='red')
summary(model4)
```
It looks like the transformations above do not help with fitting a lienar model.

#### Residuals
1. Plot a histogram of the above 4 model residuals

```{r}
options(repr.plot.width=8, repr.plot.height=8)
par(mfrow=c(2,2))

modelLog <- lm(y~log(a),data=data)
epsilonsLog <- residuals(modelLog)
hist(epsilonsLog,col='green',,main="log(a)")

modelHalf <- lm(y~I(a^0.5),data=data)
epsilonsHalf <- residuals(modelHalf)
hist(epsilonsHalf,col='green',main="a^0.50")

modelThird <- lm(y~I(a^0.33),data=data)
epsilonsThird <- residuals(modelThird)
hist(epsilonsThird,col='green',main="a^0.33")

modelQuarter <- lm(y~I(a^0.25),data=data)
epsilonsQuarter <- residuals(modelQuarter)
hist(epsilonsQuarter,col='green',main="a^0.25")
```


#### QQ
1. Plot a qqplot of the above 4 model residuals
```{r}
normalSample <- rnorm(100,0,4)
varSample <- var(normalSample)
meanSample <- mean(normalSample)
probs <- seq(1,100,1)/100 
theoreticalNormals <- qnorm(probs,meanSample,varSample)
plot(theoreticalNormals,sort(normalSample)
     ,xlab='Theoretical Normal'
     ,ylab='Empirical Distribution'
     ,tck=0.01)
data_QQ <- data.frame("sample" =sort(normalSample), "theory"=theoreticalNormals)
model = lm(sample~theory,data=data_QQ[30:70,])
abline(model,col='red')
```


```{r}
par(mfrow=c(2,2))

#log(a)
epsilonsLog = sort(epsilonsLog)
N = length(epsilonsLog)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilonsLog
      ,xlab='Theoretical Normal'
      ,ylab='Empirical Normal')
abline(lm(epsilonsLog~theoreticalNormal))
title('log(a)')

#a^0.50
epsilonsHalf = sort(epsilonsHalf)
N = length(epsilonsHalf)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)
plot(theoreticalNormal,epsilonsHalf
      ,xlab='Theoretical Normal'
      ,ylab='Empirical Normal')
abline(lm(epsilonsHalf~theoreticalNormal))
title('a^0.50')

#a^0.33
epsilonsThird = sort(epsilonsThird)
N = length(epsilonsThird)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)
plot(theoreticalNormal,epsilonsThird
      ,xlab='Theoretical Normal'
      ,ylab='Empirical Normal')
abline(lm(epsilonsThird~theoreticalNormal))
title('a^0.33')

#a^0.25
epsilonsQuarter = sort(epsilonsQuarter)
N = length(epsilonsQuarter)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)
plot(theoreticalNormal,epsilonsQuarter
      ,xlab='Theoretical Normal'
      ,ylab='Empirical Normal')
abline(lm(epsilonsQuarter~theoreticalNormal))
title('a^0.25')
```

How do the residuals and qq plot compare between your 4 different transforms?
*Thesee four plot look very alike.


#### MLR
1. Please fit a multiple linear regression model between **y** and (**x**, **a**, **b**, **c**)
```{r}
MLR <- lm(y~x+a+b+c,data=data)

plot(MLR)
summary(MLR)
```
Predicted model: y=3x-1.29a-0.22b-0.22c+6.51

2. Plot a histogram of the above 4 model residuals
```{r}
epsilons <- residuals(MLR)
hist(epsilons, col='blue',main="(x,a,b,c)")
```

3. Plot a qqplot of the above 4 model residuals
```{r}
epsilons = sort(epsilons)
N = length(epsilons)
probs = seq(1,N)/(N+1)
theoreticalNormal = qnorm(probs)

plot(theoreticalNormal,epsilons
      ,xlab='Theoretical Normal'
      ,ylab='Empirical Normal')
abline(lm(epsilons~theoreticalNormal))
title('MLR')
```


1. Does your MLR satisfy the LINE assumptions?
* No
Linearity: NO
Independent: not sure
Normality: the QQ plot looks not bad, but has some outliers
Equal Variance: the scatter plot of residuals shows the distribuiton does not have equal variance. It looks like curvy and has some extreme values. 
  
2. If not, what transforms do you recommend be applied to each explanatory variable and why?
* x: either we fit a quadratic model, or we transform y to log(y) or -1/y
  a: transformation to both x and y are necessary, but it's hard to make it linear
  b: either we fit a quadratic model, or we transform y to y^2
  c: we can try to trandform y to y^2

3. Transform the variable you feel most violates the LINE assumptions.
```{r}
plot(exp(data$x),data$y,xlab='e^x',ylab='y',tck = 0.02)
modelX <- lm(y~exp(x),data=data)
predictions <- predict.lm(modelX,data.frame(x=data$x))
lines(exp(data$x)
       ,predictions
       ,col='red')
summary(modelX)
```

4. Does transforming this variable help? why or why not?
The distribution looks more linear after this transformation.


